Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity

### 2016-06-14 analysis_parameters.tex
completed completeness and parameters sections

\textcolor{blue}{
\citet{Karimiparidari2013} applied the Maximum Curvature (MAXC) technique \citep{Wyss1999, Wiemer2000} by ZMap \citep{Wiemer2001} to calculate the level of completeness of instrumental part of the catalog.  Following the \citet{Karimiparidari2013}, we assume the catalog is complete for earthquakes with magnitude 4.5, 4.4, 4.5, 4.5, and 4.4 for Azerbaijan, Alborz,  Kopeh-Dagh, Central Iran, and Zagros tectonic seismic regions, respectively. For the uniform model we used 4.5 as a completeness magnitude. In this study we use those magnitudes of completeness as the magnitude threshold in the calculation of the seismicity parameters. We also used the seismicity parameters of \citet{Karimiparidari2013} as the priory information in HA3 code. The updated values are displayed in Table \ref{tab:b_value}.  Annual occurrence rate of the earthquake for each region is shown in Fig.~\ref{fig:annual_m}.
}

\subsection{Catalog Completness}

For the smoothed seismicity method, completeness of each magnitude in the catalog is an important factor.  \citet{Frankel1995} plotted the cumulative number of events against time for different regions. He assumes that from the point that the line become linear, the catalog is complete. This approach is similar to the MAXC method of ZMAP software \citep{Wiemer2001}, where the completeness treshold is the maximum value of the first derivative of the frequency-magnitude curve. Using the cumulative frequency-magnitude distribution of \citet{Gutenberg1944} and also frequency magnitude distribution approach in software ZMAP, \citet{Zare2014} reported the catalog completeness for the study regions. World widely large earthquakes were routinely located after increasing number of seismic stations establishment in the early 1900s \citep{Shearer2009}. Up to 1961 these data formed the early instrumental period. In 1961 the Worldwide Standardized Seismograph Network (WWSSN) was stablished. The record of these seismographs considerably improved the seismic catalogs in different part of the world \citep{Shearer2009}. Fig.~\ref{fig:completness_scatter} shows the magnitude distribution of events with respect to the time of occurrence of the events.

@@

According to   \citet{Frankel1995},  we made plots of the cumulative number of events against time for different regions' catalog. We pick the Mw 0.5 increments in magnitude to be able to compare the results with \citet{Zare2014}. In order to be able to compare the results with \citet{Zare2014} we merge the data of Azerbaijan and Alborz seismic regions.  Fig.~\ref{fig:completness_compare_zare_2014_Az_Al} shows the completeness of catalog for earthquakes with different magnitude range in Azerbaijan-Alborz region. According to this figure, midrange magnitudes ($ 4 < Mw < 6 $) fairly obey the network developments in 1900 and 1961. The completeness thresholds for each magnitude range which are reported by \citet{Zare2014} are shown by dashed green line. In this study we follow the \citet{Frankel1995} approach to determine the completeness of each region. Even though the approach used by \citet{Frankel1995} will lead to the conservative results, we make sure that we don't use a period of time without knowing the complete number of events. Determining the completeness of the catalog is very sensitive to the data. Converting earthquake magnitudes from different scales to moment magnitude obviously has some error. Having broader range of magnitude will help to minimize these sort of error. In this study we consider the magnitude intervals for completeness study as $Mw = 1$.

@@

\noindent
Fig.~\ref{fig:comp_test_all_mag} illustrates the completeness of data for each magnitude threshold in five different regions. A uniform rise of cumulative number of earthquake in each magnitude range, defines the threshold for catalog completeness. We define the completeness of each magnitude range at a time which the cumulative number of earthquake increase linearly with time. We assume the catalog for earthquakes with magnitude greater than  7 ($M_w > 7$) is complete from the first historical earthquake report.  Determining the completeness of the catalog for $ 6 < Mw \leq 7 $ needs precise observation of the catalog. In Azerbaijan tectonic seismic region after 1045  $Mw = 6.2$ earthquake the catalog seems complete up to 1318. However, between 1318 and 1581 (263 years) only one earthquake with magnitude $Mw > 6$ is reported (1440,  $Mw = 6.2$ ). Considering the activity of the Azerbaijan region, in this study we assume that the catalog is complete for earthquake with magnitude $ 6 < Mw \leq 7 $ after 1581. Similar situation happened at the period of 1715 - 1833. It worth to mention that in this case, the time period is shorter than before (118 years), meanwhile in this period of time two earthquakes with magnitude $Mw>7$ occurred in the Azerbaijan region (1721 $Mw=7.7$, 1780 $Mw=7.6$). The catalog may or may not be completed after 1045, however in this study we consider 1581 as a completeness threshold for magnitude  $ 6 < Mw \leq 7 $. Considering completeness of the catalog for Azerbaijan region from 1581 will result in conservative PGA for the region in comparison with the model with threshold of 1045. We picked the completeness threshold for other regions and other magnitude ranges according as a year that the cumulative number of earthquakes increment uniformly. Table ~\ref{tab:completeness} shows the summary of values used in this study.

>>>>

% \textcolor{blue}{
% \citet{Karimiparidari2013} applied the Maximum Curvature (MAXC) technique \citep{Wyss1999, Wiemer2000} by ZMap \citep{Wiemer2001} to calculate the level of completeness of instrumental part of the catalog.  Following the \citet{Karimiparidari2013}, we assume the catalog is complete for earthquakes with magnitude 4.5, 4.4, 4.5, 4.5, and 4.4 for Azerbaijan, Alborz,  Kopeh-Dagh, Central Iran, and Zagros tectonic seismic regions, respectively. For the uniform model we used 4.5 as a completeness magnitude. In this study we use those magnitudes of completeness as the magnitude threshold in the calculation of the seismicity parameters. We also used the seismicity parameters of \citet{Karimiparidari2013} as the priory information in HA3 code. The updated values are displayed in Table \ref{tab:b_value}.  Annual occurrence rate of the earthquake for each region is shown in Fig.~\ref{fig:annual_m}.
% }

% \subsection{Catalog Completness}

% For the smoothed seismicity method, completeness of each magnitude in the catalog is an important factor.  \citet{Frankel1995} plotted the cumulative number of events against time for different regions. He assumes that from the point that the line become linear, the catalog is complete. This approach is similar to the MAXC method of ZMAP software \citep{Wiemer2001}, where the completeness treshold is the maximum value of the first derivative of the frequency-magnitude curve. Using the cumulative frequency-magnitude distribution of \citet{Gutenberg1944} and also frequency magnitude distribution approach in software ZMAP, \citet{Zare2014} reported the catalog completeness for the study regions. World widely large earthquakes were routinely located after increasing number of seismic stations establishment in the early 1900s \citep{Shearer2009}. Up to 1961 these data formed the early instrumental period. In 1961 the Worldwide Standardized Seismograph Network (WWSSN) was stablished. The record of these seismographs considerably improved the seismic catalogs in different part of the world \citep{Shearer2009}. Fig.~\ref{fig:completness_scatter} shows the magnitude distribution of events with respect to the time of occurrence of the events.

@@

@@

% \noindent
% Fig.~\ref{fig:comp_test_all_mag} illustrates the completeness of data for each magnitude threshold in five different regions. A uniform rise of cumulative number of earthquake in each magnitude range, defines the threshold for catalog completeness. We define the completeness of each magnitude range at a time which the cumulative number of earthquake increase linearly with time. We assume the catalog for earthquakes with magnitude greater than  7 ($M_w > 7$) is complete from the first historical earthquake report.  Determining the completeness of the catalog for $ 6 < Mw \leq 7 $ needs precise observation of the catalog. In Azerbaijan tectonic seismic region after 1045  $Mw = 6.2$ earthquake the catalog seems complete up to 1318. However, between 1318 and 1581 (263 years) only one earthquake with magnitude $Mw > 6$ is reported (1440,  $Mw = 6.2$ ). Considering the activity of the Azerbaijan region, in this study we assume that the catalog is complete for earthquake with magnitude $ 6 < Mw \leq 7 $ after 1581. Similar situation happened at the period of 1715 - 1833. It worth to mention that in this case, the time period is shorter than before (118 years), meanwhile in this period of time two earthquakes with magnitude $Mw>7$ occurred in the Azerbaijan region (1721 $Mw=7.7$, 1780 $Mw=7.6$). The catalog may or may not be completed after 1045, however in this study we consider 1581 as a completeness threshold for magnitude  $ 6 < Mw \leq 7 $. Considering completeness of the catalog for Azerbaijan region from 1581 will result in conservative PGA for the region in comparison with the model with threshold of 1045. We picked the completeness threshold for other regions and other magnitude ranges according as a year that the cumulative number of earthquakes increment uniformly. Table ~\ref{tab:completeness} shows the summary of values used in this study.

### 2016-06-14 completeness.tex
completed completeness and parameters sections

@@

An important aspect of seismic catalogs is their completeness.  Completeness is is measured both in terms of the year of completeness and the completeness magnitude. Both are used in the seismic analysis process. The year of completeness is the point in time at which catalog data exhibits a steady rate of event occurrence. This is measured for different magnitude ranges. A simple approach used by \citet{Frankel1995} and others is to identify the year at which a plot of the cumulative number of events with respect to time becomes (or at least, resembles) a straight line.

The completeness magnitude, on the other hand, is the minimum magnitude at which the seismic catalog of a region can be considered complete within a certain time period. While the year of completeness can be determined by inspection of the data, the completeness magnitude requires a numerical analysis of the catalog. A common approach is to identify the magnitude associated with the maximum value of the first derivative of the frequency-magnitude curve. This method is widely used as implemented in the ZMAP software package for seismic hazard analysis developed by \citet{Wiemer2001}.

One of the difficulties in determining completeness is the availability of data for a long-enough period of time for all magnitude ranges. In this study, we determined the year of completeness for magnitude ranges 3--4, 4--5, 5--6 and 6--7 by inspection, following the approach used by \citet{Frankel1995}. This is illustrated in Fig.~\ref{fig:completeness}, where we show the cumulative number of events with time for each of the seismic zones and magnitude ranges, and indicate our choice of completeness year thresholds. For events $M_w>7$, in which case there are no sufficient data to determine a proper threshold, we assumed the catalog to be complete from the earliest time at which there is record of large magnitude earthquakes.

The chosen years of completeness are compiled in Table \ref{tab:completeness}. Also shown in this table are the values of completeness magnitude for each seismic zone and for the uniform model of the entire region of interest.

... \citet{Zare} provide the years for catalog completeness in their the original catalog for the different regions in their study on the seismicity of the Middle East. They used the

@@

\begin{figure*}[t]
    \centering
    \includegraphics[scale=0.28]{figures/pdf/figure-05.pdf} 
    \caption{Pending.}
    \label{fig:completeness}
\end{figure*}

\begin{table}[t]
    \centering
    \caption{Year thresholds and magnitude completeness for the seismic zones within the region of interest in this study.}
    \begin{tabular}{@{\hspace{0.2ex}}lccccrc@{\hspace{0.2ex}}}
        \cline{2-7}                                              \\[-1.6ex]
                        & 3--4 & 4--5 & 5--6 & 6--7 & 
                                  \multicolumn{1}{c}{$>7$} &$M_c$\\[0.6ex]
        \hline                                                   \\[-1.6ex]
        Azerbaijan      & 1961 & 1961 & 1900 & 1581 & 1042 & 4.5 \\
        Alborz          & 2005 & 1950 & 1900 & 1664 &--401 & 4.4 \\
        Kopeh Dagh      & 2005 & 1950 & 1900 & 1673 &    9 & 4.5 \\
        Zagros          & 2005 & 1961 & 1900 & 1853 & 1439 & 4.4 \\
        Central-East    & 2005 & 1961 & 1854 & 1837 &  762 & 4.5 \\
        Uniform Model   & 2005 & 1961 & 1900 & 1778 &--401 & 4.5 \\[0.5ex]
        \hline 
    \end{tabular}
    \label{tab:completeness} 
\end{table}

>>>>

\begin{figure*}[t]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/pdf/figure-05.pdf} 
    \caption{Normalized cumulative number of events and completeness year thresholds for each of the seismic zones considered within the region of interest, and for magnitude intervals of $\Delta M = 1$ between $M_w$ 3 and 6. The year thresholds are indicated by a triangle mark at the top and the vertical dashed lines in the background. Normalization values for each curve are indicated on the right margins and distinguished by color as indicated in the legend.}
    \label{fig:completeness}
\end{figure*}

@@

An important aspect of seismic catalogs is their completeness. This characteristic is measured in terms of the year of completeness and the completeness magnitude, which are used as parameters in the seismic analysis process. In simple terms, the year of completeness is the point in time at which catalog data exhibits a steady rate of event occurrence or record accumulation. This point is typically identified for different magnitude ranges. A simple approach used by \citet{Frankel1995} and others is to select the year at which a plot of the cumulative number of events with respect to time becomes (or starts to resemble) a straight line.

On the other hand, The completeness magnitude, denoted as $M_c$, is the minimum magnitude at which the seismic catalog of a region can be considered complete within a certain time period. While the year of completeness can be determined by inspection of the data, the completeness magnitude often requires a numerical analysis of the catalog. A common approach consists on identifying the magnitude associated with the maximum value of the first derivative of the frequency-magnitude curve. This method is widely used and has been implemented in the ZMAP software package for seismic hazard analysis developed by \citet{Wiemer2001}. \citet{Zare2014}, for instance, used this method to identify $M_c$ values for the different regions of the Middle East.

One of the difficulties in determining completeness is the availability of data for a period of time long enough for all magnitude ranges, so that selection of completeness parameters is reliable. In this study, we determined the year of completeness for magnitude ranges 3--4, 4--5, 5--6 and 6--7 by inspection, following the approach used by \citet{Frankel1995}. This is illustrated in Fig.~\ref{fig:completeness}, where we show the normalized cumulative number of events with time for each of the seismic zones and magnitude ranges, and indicate our choice of the completeness year thresholds in each case. For events $M_w>7$, in which case there are no sufficient data to determine a proper threshold, we assumed the catalog to be complete from the earliest time at which there is record of large magnitude earthquakes.

Admittedly, the choice of the threshold years is highly subjective. We should note then, that in the process of choosing the points indicated in the plots shown in Fig.~\ref{fig:completeness}, we considered additional contributing information such as the history of instrumentation. It is well known, for instance, that the number of recorded earthquakes increased considerably after the first deployment of seismic instruments in the early 1900s, and later with the establishment of the Worldwide Standardized Seismograph Network in 1961. We also compared our choices with the year thresholds reported by \citet{Zare2014}, which we found to be consistently earlier than our preferred years of completeness. In the end, we proceeded with our selection, which leads to a more conservative estimation of hazard.

For the case of the completeness magnitude $M_c$, we adopted the values reported by \citet{Karimiparidari2013}. Our choice the $M_c$ values by \citet{Karimiparidari2013} over those of \citet{Zare2014} is based on the fact that the former distinguishes Azerbaijan and Alborz as individual seismic zones---as we do---as opposed to the approach followed by the latter, in which these two are a single, larger seismic zone. Both the chosen years of completeness and the values of completeness magnitude for each seismic zone and for the uniform model of the entire region of interest are summarized in Table \ref{tab:completeness}.

\begin{table}%[t]
    \centering
    \caption{Year thresholds and magnitude completeness for the seismic zones within the region of interest in this study.}
    \begin{tabular}{@{\hspace{0.2ex}}lccccrc@{\hspace{0.2ex}}}
        \cline{2-7}                                              \\[-1.6ex]
                        & 3--4 & 4--5 & 5--6 & 6--7 & 
                                  \multicolumn{1}{c}{$>7$} &$M_c$\\[0.6ex]
        \hline                                                   \\[-1.6ex]
        Azerbaijan      & 1961 & 1961 & 1900 & 1581 & 1042 & 4.5 \\
        Alborz          & 2005 & 1950 & 1900 & 1664 &--401 & 4.4 \\
        Kopeh Dagh      & 2005 & 1950 & 1900 & 1673 &    9 & 4.5 \\
        Zagros          & 2005 & 1961 & 1900 & 1853 & 1439 & 4.4 \\
        Central-East    & 2005 & 1961 & 1854 & 1837 &  762 & 4.5 \\
        Uniform Model   & 2005 & 1961 & 1900 & 1778 &--401 & 4.5 \\[0.5ex]
        \hline 
    \end{tabular}
    \label{tab:completeness} 
\end{table}

% *********************************************************************************************************************
% OLD STUFF
% *********************************************************************************************************************

@@

### 2016-06-14 parameters.tex
completed completeness and parameters sections

Following the general approach described in the previous section and using the infrmation collected about the seismic catalog and its completeness, we computed the seismicity parameters $a$ and $b$, and the maximum magnitude $M_{\max}$ using the seismic hazard assessment software HA3 (version 3). This software implements the procedures developed by \citet{Kijko_1989_BSSA, Kijko_1992_BSSA}, and \citet{Kijko_2004_PAG}. HA3 is widely used for this type of computations and is regarded as a reliable tool in seismic hazard analysis \citep[see, for instance,][]{}\textcolor{red}{[references needed]}.

As mentioned before, the $a$-value is estimated based on the smoothed number of events ($M \geq 3$) per grid cell, and thus is variable throughout our region of interest. On the other hand, we consider the $b$-value and $M_{\max}$ to be constant within a given seismic zone. In this study we compute values for two different models. In the first model we obtain these parameters independently for each one of the seismic zones influencing the seismic hazard of northern Iran. That means we obtain distinct values of $b$ and $M_{\max}$ for Azerbaijan, Alborz, and Kopeh Dagh, which are our main seismic zones in northern Iran, as well as for the portions of the Zagros and Central-East seismic zones within our region of interest (see Fig.~\ref{fig:iran} for reference). In the second model we consider the whole region of interest as a uniform seismic zone. Admittedly, the second model is somewhat unrealistic if considered from the point of view that different seismic zones have different seismicity parameters. We thought, however, that computing results for a uniform model was useful for the analysis of results and comparisons with other studies.

Table \ref{tab:params} shows the parameters $b$ and $M_{\max}$ obtained using HA3 for the five seismic zones in our region of interest and for the uniform regional model, along with the observed maximum magnitude, $M_{\max}^{\mathrm{obs}}$. Fig.~\ref{fig:annualp} shows the obtained annual probability of exceedance as a function of earthquake magnitude.

>>>>

Following the general approach described in the previous sections and using the information collected about the seismic catalog and its completeness, we computed the seismicity parameters $a$ and $b$, and the maximum magnitude $M_{\max}$ using the seismic hazard assessment software HA3 (version 3). This software implements the procedures developed by \citet{Kijko_1989_BSSA, Kijko_1992_BSSA}, and subsequent updates by \citet{Kijko_2004_PAG}. HA3 is widely used for this type of computations and is regarded as a reliable tool in seismic hazard analysis \citep[see, for instance,][]{}\textcolor{red}{[references needed]}.

The $a$-value is estimated based on the completeness year, the correlation distance, and the (smoothed) number of events per grid cell. We count events with magnitude $M_w > 3$ and smooth the number of events for each cell. This means each cell has a different $a$-value and thus the result is variable throughout our region of interest. On the other hand, we consider the $b$-value and $M_{\max}$ to be constant within each of the seismic zones in our study area. We compute $b$ and $M_{\max}$ for two different models using the information from the catalog for prehistorical, historical, and instrumental seismicity and the computed values of $M_c$. In the first model we obtain these parameters independently for each one of the five seismic zones influencing the hazard in northern Iran. That is, we obtain distinct values of $b$ and $M_{\max}$ for the main northern seismic zones of Azerbaijan, Alborz, and Kopeh Dagh, as well as for the portions of the Zagros and Central-East seismic zones contributing to our region of interest (see Fig.~\ref{fig:iran} for reference).

In the second model we consider the whole study area as a uniform seismic zone. Admittedly, this assumption is unrealistic if considered from the point of view that different seismic zones have different seismicity parameters. Nonetheless, computing results for this uniform model is useful for the analysis of results and comparisons with other studies. Table \ref{tab:params} shows the parameters $b$ and $M_{\max}$ obtained with HA3 for the five seismic zones and the uniform regional model, along with the observed maximum magnitude, $M_{\max}^{\mathrm{obs}}$; and Fig.~\ref{fig:annualp} shows the annual probability of exceedance as a function of earthquake magnitude.
